String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-19 2023-1-19/sections/preliminaries.tex
all review finished
some updates
remove comments

\medskip
\noindent 
\emph{Strings and languages.}
%\paragraph*{Strings and languages.}
A string over $\Sigma$ is a (possibly empty) sequence of elements from $\Sigma$,
denoted by $u, v, w, \ldots$. An empty string is denoted by $\varepsilon$. We use $\Sigma_\varepsilon$ to denote $\Sigma \cup \{\varepsilon\}$. We write $\Sigma^*$ (resp., $\Sigma^+$) for the set of all (resp. nonempty) strings over $\Sigma$.
%Let 
For a string $u$, we use $|u|$ to denote the number of letters in $u$. In particular, $|\varepsilon|=0$. 
Moreover, for $a \in \Sigma$, let $|u|_a$ denote the number of occurrences of $a$ in $u$. 
Assume that $u=a_1\cdots a_{n}$ is nonempty and $i<j \in [n]$. %Then a \emph{position} of $u$ is a number $i \in [|u|]$ (Note that the first position is $1$, instead of  0). In addition, 
We let $u[i]$ denote $a_i$ and $u[i,j]$ for the substring %of $u$ starting from $i$ and ending with $j$ (i.e., 
$a_i\cdots a_j$. 
%\tl{check later for consistency}\zhilin{the indices start from 0, to be consistent with the semantics of substring and indexof}\zhilin{i do use it in section 4}
%
Let $u, v$ be two strings. We use $u \cdot v$ to denote the \emph{concatenation} of $u$ and $v$. A language $L$ over $\Sigma$ is a subset of strings.  
Let $L_1, L_2$ be two languages. Then the concatenation of $L_1$ and $L_2$, denoted by $L_1 \cdot L_2$,  is defined as $\{u \cdot v \mid u \in L_1, v \in L_2\}$. The union (resp. intersection) of $L_1$ and $L_2$, denoted by $L_1 \cup L_2$  (resp. $L_1 \cap L_2$), is defined as $\{u \mid u \in L_1 \mbox{ or } u \in L_2\}$ (resp. $\{u \mid u \in L_1 \mbox{ and } u \in L_2\}$). The complement of $L_1$, denoted by $\overline{L_1}$, is defined as $\{u \mid u \in \Sigma^*, u \not \in L_1\}$. 
The difference of $L_1$ and $L_2$, denoted by $L_1 \setminus L_2$, is defined as $L_1 \cap \overline{L_2}$. 
For a language $L$ and $n \in \Nat$, we define $L^n$ inductively as follows: $L^0= \{\varepsilon\}$, $L^{n+1} = L \cdot L^n$ for every $n \in \Nat$. 
Finally, define $L^* = \bigcup \limits_{n \in \Nat} L^n$.
%, that is, the string $w$ such that $|w|= |u| + |v|$ and for each $i \in [0, |u|-1]$, $w[i]= u[i]$, and for each $i \in [0,|v|-1]$, $w[|u|+i]=v[i]$.

\medskip
\noindent 
\emph{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). %Moreover, when $\delta$ is clear from context, we omit $\delta$ in $q \xrightarrow[\delta]{a} q'$ and write $q \xrightarrow{a} q'$. 
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

It is well-known that finite automata capture regular languages. 
Moreover, the class of regular languages is closed under union, intersection, concatenation, Kleene star, complement, and language difference \cite{HU79}. The constructions of NFA corresponding to these language operations are list in \ref{appendix:cefa}
% In the sequel, we recall the constructions of NFA corresponding to these language operations. 
% Let $\NFA_1 = (Q_1, \Sigma, \delta_1, I_1, F_1)$ and $\NFA_2 = (Q_2, \Sigma, \delta_2, I_2, F_2)$ be two NFA such that $Q_1 \cap Q_2 = \emptyset$. 
% \begin{itemize}
% \item The union of $\NFA_1$ and $\NFA_2$, denoted by $\NFA_1 \cup \NFA_2$, is defined as the tuple $(Q', \Sigma, \delta', I', F')$, where $Q'= Q_1 \cup Q_2$, $\delta' = \delta_1 \cup \delta_2$, $I' = I_1 \cup I_2$, and $F'=F_1 \cup F_2$. 
% %
% \item The intersection of $\NFA_1$ and $\NFA_2$, denoted by $\NFA_1 \cap \NFA_2$, is defined as the tuple $(Q', \Sigma, \delta', I', F')$, where $Q' = Q_1 \times Q_2$, $I' = I_1 \times I_2$, $F' = F_1 \times F_2$, and $\delta' = \{((q_1, q_2), a, (q'_1, q'_2)) \mid (q_1, a, q'_1) \in \delta_1, (q_2, a, q'_2) \in \delta_2\}$.
% %
% \item The concatenation of $\NFA_1$ and $\NFA_2$, denoted by $\NFA_1 \concat \NFA_2$, is defined as the tuple $(Q', \Sigma, \delta', I', F')$, where $Q' = Q_1 \cup Q_2$, $I' = I_1$, $\delta' = \delta_1 \cup \delta_2 \cup \{(q_1, a, q_2) \mid q_1 \in F_1, \exists q' \in I_2.\ (q', a, q_2) \in \delta_2\}$, moreover, if $I_2 \cap F_2 \neq \emptyset$, then $F'= F_1 \cup F_2$, otherwise, $F'= F_2$.
% %
% \item The Kleene star of $\NFA_1$, denoted by $\NFA_1^*$, is defined as $(Q'_1, \Sigma, \delta'_1, I'_1, F'_1)$, where $Q'_1 = Q_1 \cup \{q_0\}$ with $q_0 \not \in Q_1$, $I'_1 = \{q_0\}$, $F'_1= F_1 \cup \{q_0\}$, $\delta'_1 = \delta_1 \cup \{(q_0, a, q'_1)  \mid \exists q'_0 \in I_1.\ (q'_0, a, q'_1) \in \delta_1\} \cup \{(q_1, a, q'_1) \mid q_1 \in F_1, \exists q'_0 \in I_1.\ (q'_0, a, q'_1) \in \delta_1\}$.
% %
% \item The complement of $\NFA_1$, denoted by $\overline{\NFA_1}$, is defined as $(Q'_1, \Sigma, \delta'_1, q'_0, Q'_1 \setminus F'_1)$, where  $(Q'_1, \Sigma, \delta'_1, q'_0, F'_1)$ is the DFA obtained from $\NFA_1$ by subset construction.  
% %
% \item The language difference of $\NFA_1$ and $\NFA_2$, denoted by $\NFA_1 \setminus \NFA_2$, is defined as $\NFA_1 \cap \overline{\NFA_2}$.
% \end{itemize}

@@

%\tl{to put in a different place later} Note that our decision procedure also applies to general transducers. However, the EXPSPACE complexity bound is not because the distributive property $f^{-1}(L_1\cap L_2)= f^{-1}(L_1)\cap f^{-1}(L_2)$ for regular languages $L_1, L_2$ only holds for functional transducers $f$.

%We remark that an FT usually defines a relation.

%\smallskip

%In this paper, we consider logic involving two data types, i.e., the string and integer data types. We will use $u, v, \dots$ to denote string constants,  $c, d,\dots$ to denote integer constants, $x, y, \dots$ to denote string variables, and $i, j, \dots$ to denote integer variables.

\medskip
\noindent 
\emph{Linear integer arithmetic and Parikh images.}
We use standard existential \emph{linear integer arithmetic} (LIA) formulas, which
 typically range over $\psi, \varphi, \Phi, \alpha$. 
For a set $\mathfrak{X}$ of variables, we use $\psi/\varphi/\Phi/\alpha(\mathfrak{X})$ to denote the set of existential LIA formulas whose free variables are from $\mathfrak{X}$. 
For example, we use $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ to denote an LIA formula $\varphi$ whose free variables are from $\{\anivar_1, \cdots, \anivar_k\}$. For an LIA formula $\varphi(\vec{\anivar})$, we use $\varphi[\myvec{t}/\myvec{\anivar}]$ to denote the formula obtained by replacing (simultaneously) $\anivar_i$ with $t_i$ for every $i \in [k]$ where $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ and $\myvec{t} = (t_1, \cdots, t_k)$ are tuples of integer terms.
%For convenience, we assume a set $X$ of variables is a vector. 
%, which is essentially Presburger

@@

\begin{theorem}[\cite{SSMH04}]\label{thm-nfa-parikh}
Given an NFA $\NFA$,  an existential LIA formula $\psi_\NFA(\fZ_\Sigma)$ can be computed in linear time that defines the Parikh image of $\NFA$.
\end{theorem}
\vspace{-3mm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
A finite \emph{alphabet} $\Sigma$ is the set of all \emph{letters}. A
\emph{string} (or \emph{word}) is a finite sequence of letters from $\Sigma$. $\Sigma^*$ is the set of strings over $\Sigma$. $\epsilon$ is the empty string. $L$ is grammar. A language $\lan(L)$ is a set of words generated by $L$. $\mathbb{N}$ is the set of natural numbers and $\mathbb{Z}$ is the set of integer numbers. We use $a,
  b,\cdots$ to denote the constant letters in $\Sigma$, $u, v,\cdots$ to denote constant
string, $x, y,\cdots$ to denote variable string, $m,n,\cdots$ to
denote integer constant, and $i,j\cdots$ to denote integer variable. For vector, we use $\myvec{v}$ to denote the vector of integer constant, $m_n$ to denote the vector $(m,\cdots, m)$ with length $n$, $\myvec{v}[i]$ to denote the integer value of $\myvec{v}$ at position $i$, $\myvec{v_1}\cdot\myvec{v_2}$ to denote the concatenation of $\myvec{v_1}$ and $\myvec{v_2}$, $R$ to denote the vector of registers, $R_1 \cap R_2$ to denote the same registers in $R_1$ and $R_2$, and $|\myvec{v}|$ or $|R|$ to denote the length of the vector.
}

>>>>

\medskip
\noindent 
\emph{Strings and languages.}
A string over $\Sigma$ is a (possibly empty) sequence of elements from $\Sigma$,
denoted by $u, v, w, \ldots$. An empty string is denoted by $\varepsilon$. We use $\Sigma_\varepsilon$ to denote $\Sigma \cup \{\varepsilon\}$. We write $\Sigma^*$ (resp., $\Sigma^+$) for the set of all (resp. nonempty) strings over $\Sigma$.
%Let 
For a string $u$, we use $|u|$ to denote the number of letters in $u$. In particular, $|\varepsilon|=0$. 
Moreover, for $a \in \Sigma$, let $|u|_a$ denote the number of occurrences of $a$ in $u$. 
Assume that $u=a_1\cdots a_{n}$ is nonempty and $1\leq i<j \leq n$. 
We let $u[i]$ denote $a_i$ and $u[i,j]$ for the substring 
$a_i\cdots a_j$. 
%
Let $u, v$ be two strings. We use $u \cdot v$ to denote the \emph{concatenation} of $u$ and $v$. A language $L$ over $\Sigma$ is a subset of strings.  
Let $L_1, L_2$ be two languages. Then the concatenation of $L_1$ and $L_2$, denoted by $L_1 \cdot L_2$,  is defined as $\{u \cdot v \mid u \in L_1, v \in L_2\}$. The union (resp. intersection) of $L_1$ and $L_2$, denoted by $L_1 \cup L_2$  (resp. $L_1 \cap L_2$), is defined as $\{u \mid u \in L_1 \mbox{ or } u \in L_2\}$ (resp. $\{u \mid u \in L_1 \mbox{ and } u \in L_2\}$). The complement of $L_1$, denoted by $\overline{L_1}$, is defined as $\{u \mid u \in \Sigma^*, u \not \in L_1\}$. 
The difference of $L_1$ and $L_2$, denoted by $L_1 \setminus L_2$, is defined as $L_1 \cap \overline{L_2}$. 
For a language $L$ and $n \in \Nat$, we define $L^n$ inductively as follows: $L^0= \{\varepsilon\}$, $L^{n+1} = L \cdot L^n$ for every $n \in \Nat$. 
Finally, define $L^* = \bigcup \limits_{n \in \Nat} L^n$.

\medskip
\noindent 
\emph{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). \denghang{Note that one initial state is sufficient. A set of initial states is a technical choice to simplify the defination of union operation in Sec \ref{sec:automaton}.}
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

It is well-known that finite automata capture regular languages. 
Moreover, the class of regular languages is closed under union, intersection, concatenation, Kleene star, complement, and language difference \cite{HU79}. \sout{The constructions of NFA corresponding to these language operations are list in \ref{appendix:cefa}}

@@

\medskip
\noindent 
\emph{Linear integer arithmetic and Parikh images.}
We use standard existential \emph{linear integer arithmetic} (LIA) formulas, which
 typically range over $\psi, \varphi, \Phi, \alpha$. 
For a set $\mathfrak{X}$ of variables, we use $\psi/\varphi/\Phi/\alpha(\mathfrak{X})$ to denote the set of existential LIA formulas whose free variables are from $\mathfrak{X}$. 
For example, we use $\varphi(\vec{\anivar})$ with $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ to denote an LIA formula $\varphi$ whose free variables are from $\{\anivar_1, \cdots, \anivar_k\}$. For an LIA formula $\varphi(\vec{\anivar})$, we use $\varphi[\myvec{t}/\myvec{\anivar}]$ to denote the formula obtained by replacing (simultaneously) $\anivar_i$ with $t_i$ for every $i \in [k]$ where $\myvec{\anivar} = (\anivar_1, \cdots, \anivar_k)$ and $\myvec{t} = (t_1, \cdots, t_k)$ are tuples of integer terms.

@@

\begin{theorem}[\cite{SSMH04}]\label{thm-nfa-parikh}
Given an NFA $\NFA$,  an existential LIA formula $\psi_\NFA(\fZ_\Sigma)$ can be computed in linear time that defines the Parikh image of $\NFA$.
\end{theorem}
\vspace{-3mm}

### 2023-10-19 2023-1-19/sections/recl.tex
remove comments

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}

%\subsubsection{Syntax}
In the sequel, we define the string constraints with regex-counting and string-length functions, i.e., \textbf{RE}gex-\textbf{C}ounting \textbf{L}ogic (abbreviated as RECL). The syntax of RECL is defined by the rules in Figure~\ref{fig:syntax}, where $x$ is a string variable, $\anivar$ is an integer variable, $a$ is a character from an alphabet $\Sigma$, and $m, n$ are integer constants. 
A RECL formula $\varphi$ is a conjunction of atomic formulas of the form $x \in \regex$ or $t_1\ \op\ t_2$, where $\regex$ is a regular expression,  
$t_1$ and $t_2$ are integer terms, and $\op \in \{=, \neq, \le, \ge, <, >\}$. Atomic formulas of the form $x \in \regex$ are called \emph{regular membership} constraints, and atomic formulas of the form $t_1\ \op\ t_2$ are called \emph{length} constraints. 
%
A regular expression $\regex$ is built from $\emptyset$, the empty string $\epsilon$, and the character $a$ by using concatenation  $\cdot$, union $+$, Kleene star $^*$, intersection $\cap$, complement $\bar{\mbox{ }}$, difference $\setminus$, counting $^{\{m,n\}}$ or $^{\{m,\infty\}}$. An integer term is built from constants $n$, variables $\anivar$, and string lengths $|x|$ by operators $+$ and $-$.
% 
%is a quantifier-free formula that can be a regular membership $x\in\regex$, a quantifier-free Presburger formula $\alpha_1 \leq \alpha_2$, a negation of a formula, or a disjunction of two formulas. $\alpha$ is an integer term which can be an integer constant $m$, integer variable $i$, the length of the string term $|s|$, the minus of an integer term, and the plus of two integer terms. The regular expression $\regex$ is built on empty string $\epsilon$, constant letter $a\in \Sigma$. The supported regex operations involve concatenation $\cdot$, disjunction $+$, intersection $\times$, closure $*$, complement $C$, and bounded repetition $\{m, n\}$. We use $term(\varphi)$ to denote the set of terms and $strvar(\varphi)$ to denote the string variables occurring in $\varphi$. The \emph{literals} of $\varphi$ are $x\in \regex$ and $\alpha_1 \leq \alpha_2$. $x\in \regex$ is called \emph{regular literals} and $\alpha_1 \leq \alpha_2$ is called \emph{linear literals}.
\begin{figure}[h]
  \centering
%  \begin{tabular}{l r}
$ \begin{array}{l l l r}
    \varphi & ::= & x\in \regex \mid t_1\  \op\ t_2 \mid  \varphi \wedge \varphi                                              & \mbox{formulas}            \\
    \regex &::= & \emptyset \mid \epsilon \mid a \mid \regex\cdot \regex \mid \regex+\regex \mid \regex^* \mid \regex \cap \regex \mid \overline{\regex} \mid \regex \setminus \regex \mid \regex^{\{m,n\}} \mid \regex^{\{m,\infty\}} \mid (e) & \mbox{regexes} \\
    t &::= & n \mid \anivar \mid  |x| \mid t - t \mid t + t                                                                    & \mbox{integer terms}
    \end{array}
  $
%  \end{tabular}
  \caption{Syntax of RECL }\label{fig:syntax}
\vspace{-4mm}
\end{figure}
Moreover, for $S \subseteq \Sigma$ with $S = \{a_1, \cdots, a_k\}$, we use $S$ as an abbreviation of $a_1 + \cdots + a_k$.

%The definition of the semantics of RECL is standard. Nevertheless, for readability, we include the definition of its semantics here.

@@

Let $\varphi$ be a RECL formula and $\svars(\varphi)$ (resp. $\ivars(\varphi)$) denote the set of string (resp. integer) variables occurring in $\varphi$. 
%
Then the semantics of  $\varphi$ is defined with respect to a mapping $\theta: \svars(\varphi) \rightarrow \Sigma^* \uplus \ivars(\varphi) \rightarrow \Int$ (where $\uplus$ denotes the disjoint union). 
Note that the mapping $\theta$ can naturally extend to the set of integer terms. For instance, $\theta(|x|) =|\theta(x)|$, $\theta(t_1 + t_2) = \theta(t_1) + \theta(t_2)$. 
A  mapping $\theta$ is said to satisfy $\varphi$, denoted by $\theta \models \varphi$, if one of the following holds: 
%\begin{itemize}
%\item 
$\varphi \equiv x \in \regex$ and $\theta(x) \in \Lang(\regex)$, 
%
%\item $\varphi \equiv x \not \in \regex$ and $\theta(x) \not \in \Lang(\regex)$, 
%
%\item 
$\varphi \equiv t_1\ \op\ t_2$ and $\theta(t_1)\ \op\ \theta(t_2)$, 
%
%\item 
$\varphi \equiv \varphi_1 \wedge \varphi_2$ and $\theta \models \varphi_1$ and $\theta \models \varphi_2$.
%\end{itemize}
A RECL formula $\varphi$ is satisfiable if there is a mapping $\theta$ such as $\theta \models \varphi$. The satisfiability problem for RECL (which is abbreviated as $\reclsat$) is deciding whether a given RECL formula $\varphi$ is satisfiable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
We assume that $S$ is the set of string variables over $\Sigma^*$, and $I$ is the set of integer variables. $\eta: S\times\Sigma\rightarrow\Sigma^*$ is the interpretation on string where $\eta(c)=c$ for every letter $c\in \Sigma$. $\pi: I\rightarrow\mathbb{Z}$ is the interpretation of Presburger arithmetic. Then the semantics is given by a satisfaction relation: $\eta, \pi\models \varphi$ defined in Table \ref{tab:semantics}. We say a formula $\varphi$ is \emph{satisfiable} if a solution $(\eta, \pi)$ exists such as $\eta, \pi\models \varphi$. A formula $\varphi$ is \emph{unsatisfiable} if no solution exists.
\begin{table}[h]
  \centering
  \begin{tabular}{lcl}
    $\eta,\pi \models \varphi_1\vee \varphi_2$ & $\mathsf{iff}$ & $\eta,\pi \models \varphi_1 \text{ or } \eta,\pi \models \varphi_2$ \\
    $\eta,\pi \models \neg\varphi $            & $\mathsf{iff}$ & $\eta,\pi \not\models \varphi$                                      \\
    $\eta, \pi \models x\in \regex$            & $\mathsf{iff}$ & $\exists w \in \lan(\regex),\eta,\pi \models x = w$                 \\
    $\eta, \pi \models \alpha_1 \leq \alpha_2$ & $\mathsf{iff}$ & $\pi(\alpha_1) \leq \pi(\alpha_2) $                                 \\
  \end{tabular}
  \caption{Semantics}
  \label{tab:semantics}
\end{table}
%
In addition to the classic regex operators (union, concatenation, closure), we syntactically support intersection, complement, and repetition. As we all know, the classic regular language is closed under intersection and complement \cite{aut_hopcraft}. Furthermore, the operation \emph{repetition} $\regex\{m,n\}$ means repeating regex $\regex$ at least $m$ times and at most $n$ times. It can be syntactically rewritten by concatenation and union:
$\regex\{m,n\} \equiv \regex^m\mid\cdots\mid\regex^n$ where $\regex^k$ defines concatenate $\regex$ $k$ times ($m\leq k\leq n$). So the regular language $\lan(\regex)$ defined in Table \ref{tab:syntax} is semantically equal to the classic regular language.
}

%\end{document}

>>>>

%!TEX root = ../main.tex
In the sequel, we define the string constraints with regex-counting and string-length functions, i.e., \textbf{RE}gex-\textbf{C}ounting \textbf{L}ogic (abbreviated as RECL). The syntax of RECL is defined by the rules in Figure~\ref{fig:syntax}, where $x$ is a string variable, $\anivar$ is an integer variable, $a$ is a character from an alphabet $\Sigma$, and $m, n$ are integer constants. 
A RECL formula $\varphi$ is a conjunction of atomic formulas of the form $x \in \regex$ or $t_1\ \op\ t_2$, where $\regex$ is a regular expression,  
$t_1$ and $t_2$ are integer terms, and $\op \in \{=, \neq, \le, \ge, <, >\}$. Atomic formulas of the form $x \in \regex$ are called \emph{regular membership} constraints, and atomic formulas of the form $t_1\ \op\ t_2$ are called \emph{length} constraints. 
%
A regular expression $\regex$ is built from $\emptyset$, the empty string $\epsilon$, and the character $a$ by using concatenation  $\cdot$, union $+$, Kleene star $^*$, intersection $\cap$, complement $\bar{\mbox{ }}$, difference $\setminus$, counting $^{\{m,n\}}$ or $^{\{m,\infty\}}$. An integer term is built from constants $n$, variables $\anivar$, and string lengths $|x|$ by operators $+$ and $-$.
% 
\begin{figure}[h]
  \centering
$ \begin{array}{l l l r}
    \varphi & ::= & x\in \regex \mid t_1\  \op\ t_2 \mid  \varphi \wedge \varphi                                              & \mbox{formulas}            \\
    \regex &::= & \emptyset \mid \epsilon \mid a \mid \regex\cdot \regex \mid \regex+\regex \mid \regex^* \mid \regex \cap \regex \mid \overline{\regex} \mid \regex \setminus \regex \mid \regex^{\{m,n\}} \mid \regex^{\{m,\infty\}} \mid (e) & \mbox{regexes} \\
    t &::= & n \mid \anivar \mid  |x| \mid t - t \mid t + t                                                                    & \mbox{integer terms}
    \end{array}
  $
  \caption{Syntax of RECL }\label{fig:syntax}
\vspace{-4mm}
\end{figure}
Moreover, for $S \subseteq \Sigma$ with $S = \{a_1, \cdots, a_k\}$, we use $S$ as an abbreviation of $a_1 + \cdots + a_k$.

@@

Let $\varphi$ be a RECL formula and $\svars(\varphi)$ (resp. $\ivars(\varphi)$) denote the set of string (resp. integer) variables occurring in $\varphi$. 
%
Then the semantics of  $\varphi$ is defined with respect to a mapping $\theta: \svars(\varphi) \rightarrow \Sigma^* \uplus \ivars(\varphi) \rightarrow \Int$ (where $\uplus$ denotes the disjoint union). 
Note that the mapping $\theta$ can naturally extend to the set of integer terms. For instance, $\theta(|x|) =|\theta(x)|$, $\theta(t_1 + t_2) = \theta(t_1) + \theta(t_2)$. 
A  mapping $\theta$ is said to satisfy $\varphi$, denoted by $\theta \models \varphi$, if one of the following holds: 
$\varphi \equiv x \in \regex$ and $\theta(x) \in \Lang(\regex)$, 
%
$\varphi \equiv t_1\ \op\ t_2$ and $\theta(t_1)\ \op\ \theta(t_2)$, 
%
$\varphi \equiv \varphi_1 \wedge \varphi_2$ and $\theta \models \varphi_1$ and $\theta \models \varphi_2$.
A RECL formula $\varphi$ is satisfiable if there is a mapping $\theta$ such as $\theta \models \varphi$. The satisfiability problem for RECL (which is abbreviated as $\reclsat$) is deciding whether a given RECL formula $\varphi$ is satisfiable.
